---
title: "Trial analysis 2: reliability and data"
author: "Mike"
date: "2/19/2021"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    code_folding: hide
---

```{r setup, echo = FALSE}
suppressPackageStartupMessages(library(here))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(lme4))
suppressPackageStartupMessages(library(ggpmisc))
suppressPackageStartupMessages(library(ggrepel))
suppressPackageStartupMessages(library(ggthemes))
# remotes::install_github("jmgirard/agreement")
library(agreement)

# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed, cache = TRUE, 
                      message=FALSE, warning=FALSE, error=FALSE)

load(file = here("data","d_trial.Rds"))
```

The goal of this portion of the analysis is to derive two measures: 1) accuracy and 2) reaction time from the underlying looking data. Along the way we will consider various decisions we could make about data cleaning and inclusion. 

Let's start by thinking about what the basic curve is and how to get out various measures. This curve is just averaging across every single administration timepoint - unweighted. The idea is just to give a sense for what the average probability function is that we're trying to measure. 

```{r}
ggplot(d_trial, aes(x = t_norm, y = correct)) + 
  xlim(-2000,3500)+
  geom_smooth()
```

Seems like we want something that captures 1) the accuracy and 2) the rise coming soon after zero (RT).

We're going to try and develop measures for:

1. accuracy
2. baseline-corrected accuracy
3. RT

# Understanding ICCS

Here's the distribution of average looking across all trials.

```{r}
d_summary <- d_trial |>
  group_by(dataset_name, trial_id, dataset_id, subject_id, administration_id, 
           stimulus_id, target_label) |>
  summarise(accuracy = mean(correct[t_norm > 0], na.rm=TRUE),
            prop_data = mean(!is.na(correct[t_norm > 0]))) |>
  filter(!is.na(accuracy))

ggplot(d_summary, aes(x = prop_data, y = accuracy)) +
  geom_point(alpha = .05)
```

We're going to use ICCs to measure reliability, using McGraw & Wong (1996). It seems like we want two-way random effects, no interaction (subjects and items are meaningful). This is type "2A." We want average agreement across units.

One big decision is whether to look across stimulus items, rather than across kids. Across stimulus items returns *much* higher values. This is in part because we typically have more kids than items, and kids are sort of like "raters." 

```{r}
get_icc <- function (x, column = "accuracy", object = "stimulus") {
  if (object == "stimulus") {
    iccs <- dim_icc(x, 
                    model = "2A", 
                    type = "agreement", 
                    unit = "average",
                    object = stimulus_id, 
                    rater = administration_id,
                    trial = trial_id, 
                    score = {{column}}, 
                    bootstrap = 0)
  } else {
    iccs <- dim_icc(x, 
                    model = "2A", 
                    type = "agreement", 
                    unit = "average",
                    object = administration_id, 
                    rater = stimulus_id,
                    trial = trial_id, 
                    score = {{column}}, 
                    bootstrap = 0)
  }
  
  return(iccs$Inter_ICC)
}
```

## Swingley and Aslin

Let's look at one dataset. Here are the stimulus and administration ICCs for Swingley & Aslin (2002).

```{r}
sa <- d_summary |> 
  filter(dataset_name == "swingley_aslin_2002")

sa_cleaned <- filter(sa, 
         !(target_label %in% c("bird","duck","shoe","truck")))

ggplot(sa_cleaned, 
       aes(x = target_label, y = accuracy)) +
  geom_jitter(alpha = .5, width = .2) + 
  stat_summary(col = "red")
```

Now check ICCs.

```{r}
# disaggregated
get_icc(sa, object = "stimulus")
get_icc(sa, object = "administration")

# disaggregated and cleaned
get_icc(sa_cleaned, object = "stimulus")
get_icc(sa_cleaned, object = "administration")


# aggregated
sa_agg <- sa |>
  group_by(stimulus_id, administration_id) |>
  summarise(accuracy = mean(accuracy))
get_icc(sa_agg, object = "stimulus")
get_icc(sa_agg, object = "administration")
```

I don't understand the zero. Hypothesis - this is about not differentiating two different observations for each `stimulus_id`. 

```{r}
dim_icc(sa, 
        model = "2A", 
        type = "agreement", 
        unit = "average",
        object = administration_id, 
        rater = stimulus_id,
        trial = trial_id, 
        score = accuracy, 
        bootstrap = 1000)
```

```{r}
dim_icc(sa, 
        model = "2A", 
        type = "agreement", 
        unit = "average",
        object = stimulus_id, 
        rater = administration_id,
        trial = trial_id, 
        score = accuracy, 
        bootstrap = 1000)
```

Last question: inter vs. intra-rater reliability. 

```{r}
dim_icc(sa, 
        model = "2A", 
        type = "agreement", 
        unit = "average",
        object = administration_id, 
        rater = stimulus_id,
        trial = trial_id, 
        score = accuracy, 
        bootstrap = 0)
dim_icc(sa_agg, 
        model = "2A", 
        type = "agreement", 
        unit = "average",
        object = administration_id, 
        rater = stimulus_id,
        score = accuracy, 
        bootstrap = 0)
```
We're not actually sure how the **intra**-rater reliabilities are computed when you have only one observation per rater (shouldn't they be zero then?). But we're pretty clear we want the **inter**-rater reliabilities.

Take-homes:
* can't have multiple observations without a disambiguating trial label
* average absolute inter-rater reliability is what we want
* averaging across multiple observations increases reliabilities


## Across datasets

Note that we need to remove NaNs to make the ICCs work. 

```{r}
iccs <- d_summary |>
  group_by(dataset_name) |> 
  nest() |>
  mutate(icc_stimulus_acc = unlist(map(data, ~get_icc(.x, object = "stimulus"))),
         icc_admin_acc = unlist(map(data, ~get_icc(.x, object = "administration")))) |>
  select(-data) |>
  unnest(cols = c())

knitr::kable(iccs, digits = 2)
```

OK, to summarize, we think we understand the ICCs. They are:

* higher for administrations in datasets with lots of items (e.g., adams marchman)
* higher for stimuli in big datasets with few items (e.g., attword)
* reliably not NaN or 0 because we solved missing data (can't have) and repeated trials (need to mark trial id) issues. 


# Measure 1: Accuracies

There's a lot of missing data and a lot of "zoners" (kids who look only at one side). Zoners are not just missing data kids.

```{r}
ggplot(filter(d_summary, prop_data > .75),
       aes(x = accuracy)) + 
  geom_histogram()
```

Should we exclude data to get a more reliable measure? Here are two different decisions we could optimize:

1. exclude zoners?
2. exclude based on prop data

Let's try to figure those out. 

Try to do this programmatically across all datasets.  

```{r, error=FALSE, message=FALSE, warning=FALSE}

icc_sim <- function (zoners_included, exclude_less_than, object) 
{
  df <- d_summary |>
    filter(prop_data > exclude_less_than)
  
  # drop zoners
  if (zoners_included == FALSE) { 
    df <- filter(df, accuracy > 0, accuracy < 1) 
  }
  
  # compute ICCs
  df |> 
    group_by(dataset_name) |> 
    nest() |>
    mutate(icc = unlist(map(data, ~get_icc(., "accuracy", object)))) |>
    select(-data) |>
    unnest(cols = c()) 
}

excl_params <- expand_grid(zoners_included = c(FALSE, TRUE),
                           exclude_less_than = seq(.1, .9, .1), 
                           object = c("stimulus", "administration")) |>
  mutate(icc = pmap(list(zoners_included, exclude_less_than, object), icc_sim)) |>
  unnest(col = icc)
```

Plot resulting ICCs.

```{r}
ggplot(excl_params,
       aes(x = exclude_less_than, y = icc, col = zoners_included)) + 
  geom_jitter(width = .01, alpha = .5) + 
  geom_smooth(method = "lm") + 
  facet_wrap(~object)
```

Looks to me like excluding zoners isn't a clear win (and a loss for stimulus ICC). Further, excluding on amount of data doesn't seem to gain us reliability. 

I find this surprising and want to double check from other perspectives. 

## Window size

These simulations use ICCs as a way to understand how we summarize accuracy data. In particular, we're going to look at how ICCs change as a function of window size. 

```{r warning=FALSE, message=FALSE, error=FALSE}
icc_window_sim <- function (t_start = 0, t_end = 4000, object) 
{
  df <- d_trial |>
    filter(t_norm > t_start, t_norm < t_end) |>
    group_by(dataset_name, dataset_id, administration_id, stimulus_id, trial_id) |>
    summarise(accuracy = mean(correct[t_norm > 0], na.rm=TRUE),
              prop_data = mean(!is.na(correct[t_norm > 0])))
  
  # compute ICCs
  df |> 
    group_by(dataset_name) |> 
    nest() |>
    mutate(icc = unlist(map(data, ~get_icc(., "accuracy", object)))) |>
    select(-data) |>
    unnest(cols = c()) 
}

window_params <- expand_grid(t_start = seq(0,1750,250),
                             t_end = seq(2000,4000,250),
                             object = c("stimulus", "administration")) |>
  mutate(icc = pmap(list(t_start, t_end, object), icc_window_sim)) |>
  unnest(col = icc)

```

```{r}
ggplot(window_params, aes(x = t_start, y = icc, col = as.factor(t_end))) + 
  geom_jitter() + 
  facet_wrap(~object) + 
  geom_smooth(aes(group = as.factor(t_end)), se = FALSE)
```

Looks like for stimulus and administration you get consistent but modest gains if you take the longest window. BUT for stimuli, the early part of the trial adds reliability (probably because of bias due to stimulus-level preferences?). In contrast, for administrations, the early part of the trial is less informative. 500ms seems like a pretty good compromise. 

## Directions

Some notes from call on 2/17:

1. try to understand the # trial and # subject effects on ICCs. why the zeros? 
2. consider what happens when you look at times before 0, e.g. to identify zoners
3. windows that include this time period might have high stimulus reliability because they get bias. 
4. can you look at windows across age effects - older kids might need shorter windows