---
title: "Trial analysis 2B: curve shapes"
author: "Mike"
date: "10/5/2022"
output: 
  html_document:
  toc: true
toc_depth: 2
toc_float: true
code_folding: hide
---
  
```{r setup, echo = FALSE}
library(gamlss)
suppressPackageStartupMessages(library(here))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(lme4))
suppressPackageStartupMessages(library(ggpmisc))
suppressPackageStartupMessages(library(ggrepel))
suppressPackageStartupMessages(library(ggthemes))
# remotes::install_github("jmgirard/agreement")
library(agreement)
select <- dplyr::select


# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed, cache = TRUE, 
                      message=FALSE, warning=FALSE, error=FALSE)
options(dplyr.summarise.inform = FALSE)

load(file = here("cached_intermediates/","1_d_trial.Rds"))
```

# Prepare data


```{r}
BIN_INTERVAL <- 250
T_RANGE <- c(-500,3000)

ms <- d_trial %>%
  mutate(age_binned = cut(age, seq(0,60,12))) %>%
  filter(aoi %in% c("target","distractor"), 
         # dataset_name %in% c("perry_cowpig","mahr_coartic"), 
         t_norm >= T_RANGE[1],
         t_norm <= T_RANGE[2]) %>%
  mutate(t_window = 
           as.numeric(as.character(
             cut(t_norm, 
                          breaks = seq(T_RANGE[1],T_RANGE[2],BIN_INTERVAL), 
                          labels = seq(T_RANGE[1] + BIN_INTERVAL / 2, 
                                       T_RANGE[2] - BIN_INTERVAL / 2, 
                                       BIN_INTERVAL)))),
         age_centered = age - 36) %>%
  group_by(dataset_id, administration_id, trial_id, t_window, 
           age, age_centered, age_binned, target_label) %>%
  summarise(prop_target = round(mean(aoi=="target")), 
            num_target = sum(aoi == "target"), 
            N = length(aoi), 
            elogit = log( (num_target + .5) / (N - num_target + .5)), 
            wts = 1/(num_target + .5) + 1/(N - num_target + .5)) %>%
  filter(!is.na(t_window))
```

Distribution of data. 

```{r}
hist(ms$elogit)
```


```{r}
ggplot(ms, aes(x = age, y = elogit)) + 
  geom_jitter(alpha = .01)
```


Let's visualize the data going into our model, just for kicks. We can see why this is tricky. 

```{r}
ggplot(ms, 
       aes(x = t_window, y = elogit, col = age_binned)) + 
  stat_summary(fun.data = mean_se, geom = "pointrange") +
  stat_summary(geom = "line") +
  xlab("Time (ms)") +
  ylab("Target Looking (elogit)") 
```

# GAM models

```{r}
ms$dataset_id_f <- as.factor(ms$dataset_id)
ms$administration_id_f <- as.factor(ms$administration_id)
ms$target_label_f <- as.factor(ms$target_label)

mod <- gamlss(elogit ~ pb(t_window) * poly(age,2) + 
                random(dataset_id_f) +
                random(administration_id_f) + 
                random(target_label_f),
              sigma.formula = ~ pb(t_window),
              data = filter(ms, age <= 48))
       
       # (vocab ~ pbm(age, lambda = 10000),
       #                               sigma.formula = ~ pbm(age, lambda = 10000),
       #                               family = BE, 
       #                               control = gamlss.control(c.crit = .1),
       #                               data = data)
```

Model with fit. 

```{r}
summary(mod)
```

Plot. 

```{r}
ms_filt <- filter(ms, age <= 48)

ms_filt$fitted <- predict(mod, type = "response")
```


```{r}
ggplot(ms_filt, 
       aes(x = t_window, y = elogit, col = age_binned)) + 
  stat_summary(fun.data = mean_se, geom = "pointrange") +
  stat_summary(aes(y = fitted), geom = "line") +
  xlab("Time (ms)") +
  ylab("Target Looking (elogit)") 
```


# Orthogonal polynomials

Now we make our orthogonal polynomials with code straight from Mirman (2014). Make more than we need so we can try out different degrees. 

```{r}
POLY_DEG <- 5

ops <- poly(unique(ms$t_window), POLY_DEG)
ops_tibble <- tibble(ot1 = ops[,1], 
                     ot2 = ops[,2],
                     ot3 = ops[,3],
                     ot4 = ops[,4],
                     ot5 = ops[,5],
                     t_window = unique(ms$t_window))

ms <- left_join(ms, ops_tibble)
```
Plot these to understand them better. 

```{r}
ops_tibble %>%
  pivot_longer(starts_with("ot")) %>%
  ggplot(aes(x = t_window, y = value)) + 
  geom_line() +
  facet_wrap(~name)
```


# Models

We explore a glmer over the discretized proportions. This fits slowly and has some convergence issues even when we prune to the most minimal random effect structure. I wasn't able to get any random slopes in there in particular, which seems like a deal-breaker. So let's skip this. 

```{r, eval = FALSE}

mod_prop <- glmer(prop_target ~ (ot1 + ot2 + ot3) * age_centered +
                     (1 | dataset_id) +
                     (1 | target_label),
                  family = "binomial", 
                  data = ms)

```

The elogit is quite similar in distribution... not sure we can legitimately use a linear link, but we adopt it for now.

```{r}
elogit <- lmer(elogit ~ (ot1 + ot2 + ot3) * age_centered +
                  (1 | administration_id) +
                  (1 | dataset_id) +
                  (1 | target_label),
                weights = 1/wts, 
                data = ms)
```

Let's look at the model. 

```{r}
summary(elogit)
```

Now, let's look at model fits. 

```{r}
elogit_data = ms
elogit_data$fit <- fitted(elogit)

ggplot(elogit_data, 
       aes(x = t_window, y = elogit, col = age_binned)) + 
  stat_summary(fun.data = mean_se, geom = "pointrange") +
  stat_summary(aes(y = fit), fun = mean, geom = "line") +
  geom_hline(yintercept = 0, col = "black", lty = 2) +
  # ylim(0, 1) +
  xlab("Time (ms)") +
  ylab("Target Looking (elogit)") 
```
So x is the elogit value. 

$$x = \log(\frac{y + .5}{N - y + .5})$$

So solve for y

$$\frac{e^x(2N + 1) - 1}{2(e^x + 1)} = y  $$

The `VWpre` package has a function, but it assumes constant N. 

```{r}
elogit_to_prop <- function(x, N) {
  y = ((exp(x) * (2 * N + 1) - 1) / (2 * (exp(x) + 1))) / N
  
  return(y)
}
```

Plot in proportion space. 

```{r}
# try VWPRE
elogit_data$fitted_prop <- elogit_to_prop(elogit_data$fit, elogit_data$N)
ggplot(elogit_data, 
       aes(x = t_window, y = prop_target, col = age_binned)) + 
  stat_summary(fun.data = mean_se, geom = "pointrange") +
  stat_summary(aes(y = fitted_prop), fun = mean, geom = "line") +
  geom_hline(yintercept = .5, col = "black", lty = 2) +
  # ylim(0, 1) +
  xlab("Time (ms)") +
  ylab("Proportion Target Looking") 
```


Plot model predictions from fixed effects. **TODO: quadratic age effects**

```{r}
newdata <- left_join(expand_grid(t_window = seq(50, 1450, 100), 
                            age_centered = c(18,30, 42, 54) - 36), 
                            ops_tibble) 

newdata$pred <- predict(elogit, newdata = newdata, re.form = NA)
newdata$fitted_prop <- elogit_to_prop(newdata$pred, 4)

ggplot(newdata, 
       aes(x = t_window, y = pred, col = factor(age_centered+36))) + 
  geom_line()
```
