---
title: 'Trial analysis 3B: modeling individual RTs'
author: "Peekbank team"
date: "2022-08-16"
output: html_document
---

```{r setup, echo = FALSE, message = FALSE}
library(here)
library(tidyverse)
library(lme4)
library(ggthemes)
library(cowplot)

# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed, cache = TRUE, 
                      message=FALSE, warning=FALSE, error=FALSE)

load(file = here("data","d_rt_dt.Rds"))
```

Goal: build interesting models of RT data. 

These are D-T shifts with RTs > 250 ms. 

# Visualization

```{r}
ggplot(d_rt_dt, aes(x = rt)) + 
  geom_histogram()
```
Logs are aesthetically nicer, though distribution is not perfectly log normal still. 

```{r}
ggplot(d_rt_dt, aes(x = rt)) + 
  geom_histogram() + 
  scale_x_log10()
```
Look by dataset. 

```{r}
ggplot(d_rt_dt, aes(x = rt)) + 
  geom_histogram() + 
  scale_x_log10() + 
  facet_wrap(~dataset_name, scales = "free_y")
```
Feeling good about this as a dependent variable to model. 

Relation with age shows some bumpiness perhaps because of variation across datasets. 
```{r}
ggplot(d_rt_dt, aes(x = age, y = log_rt)) + 
  geom_jitter(alpha = .03, width = .5) + 
  geom_smooth()
```
```{r}
ggplot(d_rt_dt, aes(x = age, y = log_rt)) + 
  geom_jitter(alpha = .1, width = .5) + 
  geom_smooth(method = "loess", span = 1) + 
  facet_wrap(~dataset_name)
```

# Model development

Log and center age for fitting. 

```{r}
d_rt_dt$log_age <- log(d_rt_dt$age)
d_rt_dt$log_age_centered <- d_rt_dt$log_age - mean(d_rt_dt$log_age)
```

Need to get high-frequency words for plotting purposes.

```{r}
words <- d_rt_dt |>
  group_by(target_label) |> 
  count() 

hf_words <- words |>
  filter(n > 100)
```


First things first: age vs. log age. 

```{r}
age_mod <- lmer(log_rt ~ age + 
                  (1 | administration_id) + 
                  (1 | dataset_name) + 
                  (1 | target_label), 
                data = d_rt_dt, 
                control = lmerControl(optimizer="bobyqa"))

log_age_mod <- lmer(log_rt ~ log_age_centered + 
                  (1 | administration_id) + 
                  (1 | dataset_name) + 
                  (1 | target_label), 
                data = d_rt_dt, 
                control = lmerControl(optimizer="bobyqa"))

anova(age_mod, log_age_mod)
```

Conclusion: log age is very slightly better but not so much better given the quantity of data. 

```{r}
mod <- lmer(log_rt ~ log_age_centered + 
              (1 | administration_id) + 
              (1 | dataset_name) + 
              (log_age_centered | target_label), 
            data = d_rt_dt)
```
Having trouble figuring out what to do about convergence.

* random slopes of age by dataset have limited variance. 
* can't get any distractor effects in there. 

```{r}
summary(mod)
```

Predicted curves. 

```{r}
preds <- expand_grid(
  log_age_centered = seq(min(d_rt_dt$log_age_centered),
                max(d_rt_dt$log_age_centered), .1),
  target_label = hf_words$target_label)

preds <- preds |>
  mutate(.pred = predict(mod, 
                         type = "response",
                         re.form = ~  (log_age_centered | target_label), 
                         newdata = preds), 
         age = exp(log_age_centered + mean(d_rt_dt$log_age)))

ggplot(d_rt_dt,
       aes(x = age, y = log_rt)) + 
  geom_point(alpha = .1) + 
  geom_line(data = preds, 
            aes(x = age, y = .pred, col = target_label)) + 
  geom_label_repel(data = filter(preds, age == max(age)), 
                   aes(x = age, y = .pred, col = target_label, 
                       label = target_label), size = 3, max.overlaps = 50,
                   force = .5) + 
  xlab("Age (months)") + 
  ggthemes::theme_few()
```


# Animacy analysis

Animacy played a big role for accuracy. Does it for RT? 

```{r}

animates <- c("animal","baby","dog","bear","bird","boy", 
              "bunny","cat","chicken","cow","dog","duck","elephant",
              "fish","frog","giraffe","horse","kangaroo","lion","monkey",
              "owl","pig","puppy","sheep","teddy","tiger","whale","zebra")

d_rt_dt$animate_target <- d_rt_dt$target_label %in% animates
d_rt_dt$animate_distractor <- d_rt_dt$distractor_label %in% animates

animates_mod <- lmer(log_rt ~ log_age_centered * animate_target + 
                       log_age_centered * animate_distractor + 
              (1 | administration_id) + 
              (1 | dataset_name) + 
              (log_age_centered | target_label), 
            data = d_rt_dt)

summary(animates_mod)
```

Doesn't look like it, surprisingly. 

```{r}
preds <- expand_grid(
  animate_target = c(TRUE, FALSE), 
  animate_distractor = c(TRUE, FALSE), 
  log_age_centered = seq(min(d_rt_dt$log_age_centered),
                max(d_rt_dt$log_age_centered), .1),
  target_label = hf_words$target_label)

preds <- preds |>
  mutate(.pred = predict(animates_mod, 
                         type = "response",
                         re.form = NA, 
                         newdata = preds), 
         age = exp(log_age_centered + mean(d_rt_dt$log_age)))


ggplot(d_rt_dt,
       aes(x = age, y = log_rt)) + 
  geom_point(alpha = .1) + 
  geom_line(data = preds, 
            aes(y = .pred, col = animate_target, lty = animate_distractor), 
            size = 1.5) + 
  scale_x_continuous(breaks = c(12,24,36,48,60)) + 
  scale_color_solarized(name = "Animate target") + 
  scale_linetype(name = "Animate distractor") + 
  ylab("Trial-level log(RT)") + 
  xlab("Age (months)") + 
  ggthemes::theme_few()
```


# Export

```{r}
mod_rt <- mod
d_rt <- d_rt_dt
save(file = here("cached_intermediates","3B_mod_rt.Rds"), mod_rt)
save(file = here("cached_intermediates","3B_d_rt.Rds"), d_rt)
```


```

